---
title: "R NOTEBOOK - **Sequential Rules Analysis**"
output: html_notebook
---

We first defined the working directory, load the required library and load the transaction dataset.   
```{r}
# Set Working Directory
setwd("~/SANDY/web_analytics/1_CA1/CA1_r11/3_Sequential_Rules_Analysis")

#Load Required Libraries
library(arulesViz)
library(dplyr)
library(datetime)
library(arulesSequences)

#Load Required data 
dt1 = read.csv(file="train.csv")

```

#### The following is the first **5** rows of the dataset.
```{r}
dt1[1:5,]
```

#### The following is the **Basic Summary** of the dataset.
```{r}
summary(dt1)
```

#### Transform the dataset.
```{r}
df <- dt1 %>%
      arrange(datetime_y) %>%
      arrange(session_id) %>%
      unique() %>%
      group_by(session_id) %>%
      #summarise(cart=paste(item_id,collapse=";")) %>%
      ungroup()
df[1:5,]
```
Create additional columns for transformation. 
```{r}
df$sequenceID <- df$session_id
df$eventID    <- df$session_id
df$SIZE       <- '1'
df$items      <- df$item_id
df$items      <- as.factor(df$items)
df[1:5,]
```

#### Recode **Event Id** to numeric ascending order.
```{r}
df$eventID <- df$eventID[1] <- 1
for (i in 1:length(df$sequenceID)) {
  
  if (i == 1) {df$eventID[i] <- 1} else
  if( df$sequenceID[i-1] == df$sequenceID[i])
  {
      df$eventID[i] <-df$eventID[i-1]+1 
    
    }
}
```
```{r}
df[1:5,]
```

```{r}
df1 <- df
# Check dummy columns
df1$seq_test <- df1$sequenceID
df1$sequenceID <-df1$sequenceID[1] <- 1

df1[1:5,]
```

#### Recode **Sequence Id** to numeric ascending order.
```{r}
for (i in 1:length(df1$seq_test)) {
  
  if (i == 1) {df$sequenceID[i] <- 1} else 
  if(df1$seq_test[i-1] == df1$seq_test[i]){ df1$sequenceID[i] = df1$sequenceID[i-1] }
  else {df1$sequenceID[i] = df1$sequenceID[i-1]+1}
}
```
```{r}
df1[1:5,]
```

#### The **final sequence format** data is as follows:-
```{r}
df2           <- df1[c(5,6,7,8)]
df2$sequenceID <- as.integer(df2$sequenceID)
df2$eventID   <- as.integer(df2$eventID)
df2$SIZE      <- as.integer(df2$SIZE)
df2 <- df2[order(df2$sequenceID,df2$eventID),]
#seqchkpt1
df2[1:5,]
```

#### Export the data out as **.txt** files and re-construct the **Transaction Basket** file.
```{r}
write.table(df2, "seq_format.txt", sep=" ", row.names = FALSE, col.names = FALSE, quote = FALSE)
data <- read_baskets(con = "seq_format.txt", info = c("sequenceID","eventID","SIZE"))
```

#### Show **Transaction Object** Information
```{r}
 transactionInfo(data)
```

#### Show the **Sequences Rules**.
```{r}
as(head(data), "data.frame")
```

#### Run **CSpade Algorithm**.    
For CSAPDE algoritm you might set some lags so that you can extract rules from sequence of transactions with the lag.   
We set the minimum support of rules to **0.5%**.
```{r}
seqs <- cspade(data, parameter = list(support = 0.0005), control = list(verbose = TRUE))
```

#### View the **Sequences**.
```{r}
as(seqs,"data.frame")  # view the sequences
```

#### Convert extracted sequential rules to data frame and Filter rules with more than one sequence
```{r}
scrul.dt <- as(seqs,"data.frame")
scrul.dt$sequence <- gsub("df3\\$cart2\\=|<|>","",scrul.dt$sequence)

scrul.dt1 <- scrul.dt[count.fields(textConnection(scrul.dt$sequence),sep = ",")>1,]
scrul.dt1
```

```{r}
scrul.dt1[10,]
```
Each of unique sequences happened on the same date. For rule 10,If a customer’s first purchase is 214853102, his second purchase would be 214854840 which is frequent for around 2% of session user.

#### Induced the Sequences Rules.
```{r}
seqrules <- ruleInduction(seqs, confidence = 0.5,control = list(verbose = TRUE))
```

#### The following is the **Sequence Rules with 50% Confidences**.
```{r}
as(seqrules,"data.frame")  # view the rules
```

### **Testing Sequence Rules**    
We first defined the **Working Functions** and load the **Test** dataset
```{r}
#remove duplicate items from a basket (itemstrg)
uniqueitems <- function(itemstrg) {
  unique(as.list(strsplit(gsub(" ","",itemstrg),","))[[1]])
}
# execute ruleset using item as rule antecedent (handles single item antecedents only)
makepreds <- function(item, rulesDF) {
  antecedent = paste("<{",item,"}> =>",sep="") # NOTE: diff from assoc analysis same fn
  firingrules = rulesDF[grep(antecedent, rulesDF$rule,fixed=TRUE),1] # rules is now rule
  #gsub(" ","",toString(sub(">}","",sub(".*=> <{","",firingrules))))
  gsub(" ", "", toString(sub('\\}>', '', sub(".*=> <\\{", "", firingrules))))
}
# count how many predictions are in the basket of items already seen by that user 
# Caution : refers to "baskets" as a global
checkpreds <- function(preds, baskID) {
  plist = preds[[1]]
  blist = baskets[baskets$basketID == baskID,"items"][[1]]
  cnt = 0 
  for (p in plist) {
    if (p %in% blist) cnt = cnt+1
  }
  cnt
}
# count all predictions made
countpreds <- function(predlist) {
  len = length(predlist)
  if (len > 0 && (predlist[[1]] == "")) 0 # avoid counting an empty list
  else len
}
# Load the test data
testegs = read.csv(file="test.csv")
testegs = testegs[,c(1,3)]
colnames(testegs) <- c("basketID","items")  # set standard names, in case they are different in the data file
# Display top 5 rows
testegs[1:5,]
```

```{r}
#execute rules against test data
rulesDF = as(seqrules,"data.frame")
testegs$preds = apply(testegs,1,function(X) makepreds(X["items"], rulesDF))

# extract unique predictions for each test user
userpreds = as.data.frame(aggregate(preds ~ basketID, data = testegs, paste, collapse=","))
userpreds$preds = apply(userpreds,1,function(X) uniqueitems(X["preds"]))

# extract unique items for each test user
baskets = as.data.frame(aggregate(items ~ basketID, data = testegs, paste, collapse=","))
baskets$items = apply(baskets,1,function(X) uniqueitems(X["items"]))

#count how many unique predictions made are correct, i.e. have previously been bought (or rated highly) by the user
correctpreds = sum(apply(userpreds,1,function(X) checkpreds(X["preds"],X["basketID"])))

# count total number of unique predictions made
totalpreds = sum(apply(userpreds,1,function(X) countpreds(X["preds"][[1]]))) 
precision = correctpreds*100/totalpreds
cat("precision=", precision, "corr=",correctpreds,"total=",totalpreds)
```
